Implications of afforestation policies for global markets

### 2023-08-07 paper.Rmd
Update paper.Rmd

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020]. In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020]. In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available. This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021]. There are many reasons for this. For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets. Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland.

Land-use decisions in one countr also tend to have impacts that spill over into many others [@meyfroidt2010] [@lambin2011], and therefore, cumulative change needs to be assessed with global-scale models.  Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets. The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services. While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017]. There have been attempts to use the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but the standard model does not well account for land-use changes. The Land Use and Land Cover Database within the GTAP framework, which incorporated forestry remote-sensing products, was thus created [@baldos2012]. It was strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008]. It also includes competition among different crops, grazing, and forest-based uses, but it has limitations due to its infrequent updates. The KLUM@GTAP framework integrates the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation. GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions. KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions. KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].

@@

Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

>>>>

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020]. In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020]. In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available. This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021]. There are many reasons for this. For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets. Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland. Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017]. However, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.

Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets. The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services. While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017]. Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes. As a result, the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products, was created [@baldos2012]. It was further strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008]. While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates. The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation. GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions. KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions. KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].

@@
